### 2017-02-21 docs/binary.tex
Trunk add LC emulsion document and relevant changes
git-svn-id: https://ccpforge.cse.rl.ac.uk/svn/ludwig/trunk@3079 40876441-3496-0410-a613-a2fca35b2ae2

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  binary.tex
%
%  Symmetric binary fluids and Brazovskii Smetics
%
%  Edinburgh Soft Matter and Statistical Physics Group and
%  Edinburgh Parallel Computing Centre
%
%  (c) 2016 The University of Edinburgh
%
%  Contributing authors:
%  Kevin Stratford (kevin@epcc.ed.ac.uk) and I've taken material from
%  Grace Kim's                           doctoral thesis (UoE 2009)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Binary Fluids}

@@

\includegraphics[width=0.9\textwidth]{figures/symmetric.eps}

@@

If we consider a one-dimensional case, this is
\begin{equation}
A\phi + B\phi^3 - \kappa \frac{d^2 \phi}{dx^2} = 0.
\end{equation}
Note that the conservation of order parameter means this should be a
constrained minimisation, which can be handled by adding the
a Lagrange multiplier $\Lambda = \partial f / \partial \phi |_{\phi_0}$.
The resulting ordinary differential equation has a solution
$\phi(x) = \phi^\star \tanh(x/\xi)$ where $\xi$ is identified
as the interfacial width. It can be confirmed that this is a solution
if
\begin{equation}
\xi = (-2\kappa / A)^{1/2}.
\end{equation}
The one-dimensional case can also be used to determine the interfacial
tension, which is the excess energy associated with the interface.
If we denote the interfacial tension $\sigma$, then the excess
energy is
\begin{equation}
\sigma = \int_{-\infty}^{+\infty} \mathrm{d} x
\Big[
{\textstyle \frac{1}{2}} \kappa (d_x \phi)^2 + V(\phi) - V(\phi^\star)
\Big],
\end{equation}
which can be evaluated from the equilibrium profile with the interface
placed at the origin $\phi(x) = \phi^\star\tanh(x/\xi)$, along with a number
of standard integrals, to give
\begin{equation}
\sigma =  4\kappa\phi^{\star 2}/3\xi
 = \big(-8\kappa A^3/9B^2\big)^{1/2}.
\end{equation}
Further discussion of the practical choice of the free energy parameters,
and their
impact on interfacial width and tension is deferred until LATER.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  binary.tex
%
%  Symmetric binary fluids and Brazovskii Smetics
%
%  Edinburgh Soft Matter and Statistical Physics Group and
%  Edinburgh Parallel Computing Centre
%
%  (c) 2016-2017 The University of Edinburgh
%
%  Contributing authors:
%  Kevin Stratford (kevin@epcc.ed.ac.uk) and I've taken material from
%  Grace Kim's                           doctoral thesis (UoE 2009)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Binary Fluids}
\label{section-binary}

@@

\includegraphics[width=0.9\textwidth]{figures/symmetric}

@@

If we consider a one-dimensional case, this is
\begin{equation}
A\phi + B\phi^3 - \kappa \frac{d^2 \phi}{dx^2} = 0.
\end{equation}
Note that the conservation of order parameter means this should be a
constrained minimisation, which can be handled by adding the
a Lagrange multiplier $\Lambda = \partial f / \partial \phi |_{\phi_0}$.
The resulting ordinary differential equation has a solution
$\phi(x) = \phi^\star \tanh(x/\xi)$ where $\xi$ is identified
as the interfacial width. It can be confirmed that this is a solution
if
\begin{equation}
\xi = (-2\kappa / A)^{1/2}.
\end{equation}
The one-dimensional case can also be used to determine the interfacial
tension, which is the excess energy associated with the interface.
If we denote the interfacial tension $\sigma$, then the excess
energy is
\begin{equation}
\sigma = \int_{-\infty}^{+\infty} \mathrm{d} x
\Big[
{\textstyle \frac{1}{2}} \kappa (d_x \phi)^2 + V(\phi) - V(\phi^\star)
\Big],
\end{equation}
which can be evaluated from the equilibrium profile with the interface
placed at the origin $\phi(x) = \phi^\star\tanh(x/\xi)$, along with a number
of standard integrals, to give
\begin{equation}
\sigma =  4\kappa\phi^{\star 2}/3\xi
 = \big(-8\kappa A^3/9B^2\big)^{1/2}.
\end{equation}
Further discussion of the practical choice of the free energy parameters,
and their
impact on interfacial width and tension is deferred until
Section~\ref{section-examples}.

\subsubsection{The free energy density in practice}

Writing the free energy density in full, we have
\begin{equation}
\label{eq-binary-fed}
f(\phi, \partial_\alpha \phi) =
{\textstyle\frac{1}{2}} A \phi^2 +
{\textstyle\frac{1}{4}} B \phi^4 +
{\textstyle\frac{1}{2}} \kappa (\partial_\alpha \phi)^2.
\end{equation}
Note that some authors define $A$ to be a positive constant in which case
a negtive sign appears in the quadratic term above. We retain $A$ negative
to describe the separated case.

### 2017-02-21 docs/lb.tex
Trunk add LC emulsion document and relevant changes
git-svn-id: https://ccpforge.cse.rl.ac.uk/svn/ludwig/trunk@3079 40876441-3496-0410-a613-a2fca35b2ae2

We review here the lattice Boltzmann method applied to a simple
Newtonian fluid with particular emphsis on the relevant
implementation in \textit{Ludwig}.

@@

In lattice Boltzmann, the density and velocity of the continuum fluid
are complemented by the  distribution function
$f_i(\mathbf{r}; t)$ defined with reference to the
discrete velocity space $c_{i\alpha}$.
It is possible to relate the hydrodynamic quantities to the distribution
function via its moments, that is
\begin{equation}
\rho(\mathbf{r};t) = \sum_i f_i(\mathbf{r};t),  \quad
\rho u_\alpha(\mathbf{r};t) = \sum_i f_i(\mathbf{r};t) c_{i\alpha},  \quad
\Pi_{\alpha\beta}(\mathbf{r};t) =
\sum_i f_i(\mathbf{r};t) c_{i\alpha} c_{i\beta}.
\label{lb-f-moments}
\end{equation}
Here, the index of the summation is over the number of discrete
velocities used as the basis, a number which will be denoted
$N_\mathrm{vel}$. For example, in three dimensions
$N_\mathrm{vel}$ is often 19 and the basis is referred to as D3Q19.

The number of moments, or modes, supported by a velocity
set is exactly $N_\mathrm{vel}$, and these can be written in general as
\begin{equation}
M^a(\mathbf{r};t) = \sum_i m_i^a f_i(\mathbf{r};t),
\end{equation}
where the $m_i$ are the eigenvectors of the collision matrix in the LBE.
For example, in the case of the density, all the 
$m_i^a = 1$ and the mode $M^a$ is the density $\rho = \sum_i f_i$. Note
that the number of modes supported by a given basis will generally excceed the
number of hydrodynamic modes; the excess modes have no direct physical
interpretation and are variously referred to as non-hydrodynamic, kinetic,
or ghost, modes.
The ghost modes take no part in bulk hydrodynamics, but may become important
in other contexts, such as thermal fluctuations and near boundaries.
The distribution function can be related to the modes
$M^a(\mathbf{r};t)$ via
\begin{equation}
f_i(\mathbf{r};t) = w_i \sum_a m_i^a N^a M^a(\mathbf{r};t).
\end{equation}
In this equation, $w_i$ are the standard LB weights appearing in the
equilibrium distribution function, while the $N^a$ are a per-mode
normalising factor uniquely determined by the orthogonality condition
\begin{equation}
N^a \sum_i w_i m_i^a m_i^b = \delta_{ab}.
\end{equation}
Writing the basis this way has the advantage that the equilibrium
distribution projects directly into the hydrodynamic modes only.
Putting it another way, we may write
\begin{equation}
f_i^\mathrm{eq} = w_i \big(\rho + \rho c_{i\alpha}u_\alpha / c_s^2
+ (c_{i\alpha} c_{i\beta} - c_s^2\delta_{\alpha\beta})
(\Pi_{\alpha\beta}^\mathrm{eq} - p\delta_{\alpha\beta})/2c_s^4 \big)
\end{equation}
where only (equilibrium) hydrodynamic quantities appear on the right hand side.

@@

\begin{table}[t]
\begin{center}
\begin{tabular}{|l|r|rrrrrrrrr|r|l|}
\hline\hline
$M^a$ & $p$ & \multicolumn{9}{c|}{$m_i^a$} & $N^a$  &\\
\hline
$\rho$ & - & 1 &  1 &  1 &  1 &  1 &  1 &  1 &   1 &  1 & 1 &$\mathbf{1}$ \\
\hline
$\rho c_{ix}$ & - & 0 &  1 &  1 & 1 & 0 &  0 & -1 &  1 & -1 & 3 & $c_{ix}$ \\
\hline
$\rho c_{iy}$ & - & 0 & 1 &  0 &  -1 &  1 &  -1 & 1 & 0 & -1 & 3  &$c_{iy}$ \\
\hline
$Q_{xx}$ & 1/3 & -1 &  2 &  2 & 2 & -1 & -1 & 2 & 2 & 2 & 9/2 
& $c_{ix} c_{ix} - c_s^2$ \\
\hline
$Q_{xy}$ & - & 0 &  1 & 0 & -1 & 0 & 0 & -1 & 0 & 1 & 9 & $c_{ix} c_{iy}$ \\
\hline
$Q_{yy}$ & 1/3 & -1 &  2 & -1 & 2 & 2 & 2 & 2 & -1 & 2 & 9/2
& $c_{iy} c_{iy} - c_s^2$ \\
\hline\hline
$\chi^1$ & - &  1 & 4 & -2 & 4 & -2 & -2 & 4 & -2 & 4 & 1/4 & $\chi^1$ \\
\hline
$J_{ix}$ & - & 0 &  4 & -2 & 4 & 0 & 0 & -4 & -2 & -4 & 3/8
& $\chi^1 \rho c_{ix}$\\
\hline
$J_{iy}$ & - & 0 & 4 & 0 & -4 & -2 & 2 & 4 & 0 & -4 & 3/8
& $\chi^1 \rho c_{iy}$\\
\hline\hline
$w_i$ & 1/36 & 16 & 1 & 4 & 1 & 4 & 4 & 1 & 4 & 1 & & $w_i$\\
\hline\hline
\end{tabular}
\end{center}
\caption{Table showing the details of the basis used for the D2Q9 model
in two dimensions. The nine modes $M^a$ include six hydrodynamic modes,
one scalar kinetic mode $\chi^1$, and one vector kinetic mode $J_{i\alpha}$.
The weights in the equilibrium distribution function are $w_i$ and the
normaliser for each mode is $N^a$. The eigenvectors of the collision
matrix are the columns of the transformation matrix $m^a_i$. The prefactor
$p$ (where present) multiplies all the elements to the right in that row.}
\label{table-d2q9-spec}
\end{table}

@@

\caption{Table showing the details of the basis used for the D3Q15 model
in three dimensions. The fifteen modes $M^a$ include two scalar kinetic
modes $\chi^1$ and $\chi^3$, and one vector kinetic mode $J_{i\alpha}$.
The weights in the equilibrium distribution are $w_i$ and the normaliser
for each mode is $N^a$. The eigenvectors of the collision matrix are the
columns of the transformation matrix $m_i^a$. The prefactor $p$ simply
multiplies all elements of $m_i^a$ in that row as a convenience.
\label{table-d3q15-spec}
}
\end{table}

@@

\begin{table}[t]
\centering
\tabcolsep=4pt
\begin{tabular}{|l||r|rrrrrr|rrrr|rrrr|rrrr|r||}
\hline\hline
$M^a$ & \multicolumn{19}{c||}{$m_i^a$} & $N^a$\\
\hline
$\rho $ & 1 &  1 &  1 &  1 &  1 &  1 &  1 & 
         1 &  1 &  1 &   1 &  1 &  1 &  1 & 1 & 1 & 1 & 1 & 1
& 1\\
\hline
$\rho c_{ix}$ & 0 &  1 &  -1 &  0 &  0 &  0 &  0 & 
         1 &  1 &  -1 &   -1 &  1 &  1 &  -1 & -1 & 0 & 0 & 0 & 0
& 3 \\
\hline
$\rho c_{iy}$ & 0 &  0 &  0 &  1 &  -1 &  0 &  0 & 
         1 &  -1 &  1 &   -1 &  0 &  0 &  0 & 0 & 1 & 1 & -1 & -1
& 3\\
\hline
$\rho c_{iz}$ & 0 &  0 &  0 &  0 &  0 &  1 &  -1 & 
         0 &  0 &  0 &   0 &  1 &  -1 &  1 & -1 & 1 & -1 & 1 & -1
& 3\\
\hline
$Q_{ixx}$ & -1 &  2 &  2 &  -1&  -1 &  -1 &  -1 & 
         2 &  2 &  2 &   2 &  2 &  2 &  2 & 2 & -1 & -1 & -1 & -1
& 9/2\\
\hline
$Q_{iyy}$ & -1 &  -1 &  -1 &  2&  2 &  -1 &  -1 & 
         2 &  2 &  2 &   2 &  -1 &  -1 &  -1 & -1 & 2 & 2 & 2 & 2
& 9/2\\
\hline
$Q_{izz}$ & -1 &  -1 &  -1 &  -1&  -1 &  2 &  2 & 
         -1 &  -1 &  -1 &   -1 &  2 &  2 & 2 & 2 & 2 & 2 & 2 & 2
& 9/2\\
\hline
$Q_{ixy}$ & 0 &  0 &  0 &  0&  0 &  0 &  0 & 
          1 &  -1 &  -1 &    1 &  0 &  0 & 0 & 0 & 0 & 0 & 0 & 0
& 9\\
\hline
$Q_{ixz}$ & 0 &  0 &  0 &  0&  0 &  0 &  0 & 
          0 &   0 &   0 &   0 &  1 & -1 & -1 & 1 & 0 & 0 & 0 & 0
& 9\\
\hline
$Q_{iyz}$ & 0 &  0 &  0 &  0&  0 &  0 &  0 & 
          0 &   0 &   0 &   0 &  0 & 0 & 0 & 0 & 1 & -1 & -1 & 1
& 9\\
\hline\hline
$\chi^1$ & 0 &  1 &  1 &  1 &  1 &  -2 &  -2 & 
         -2 &  -2 &  -2 &  -2 &  1 &  1 & 1 & 1 & 1 & 1 & 1 & 1
& 3/4\\
\hline
$\chi^1 \rho c_{ix}$ & 0 &  1 &  -1 &  0&  0 &  0 &  0 & 
         -2 &  -2 &  2 &  2 &  1 &  1 & -1 & -1 & 0 & 0 & 0 & 0
& 3/2\\
\hline
$\chi^1 \rho c_{iy}$ & 0 &  0 &  0 &  1&  -1 &  0 &  0 & 
         -2 &  2 &  -2 &  2 &  0 &  0 & 0 & 0 & 1 & 1 & -1 & -1
& 3/2\\
\hline
$\chi^1 \rho c_{iz}$ & 0 &  0 &  0 &  0&  0 &  -2 &  2 & 
         0 &  0 &  0 &  0 &  1 &  -1 & 1 & -1 & 1 & -1 & 1 & -1
& 3/2\\
\hline
$\chi^2$ & 0 &  1 &  1 &  -1&  -1 &  0 &  0 & 
         0 &  0 &  0 &  0 &  -1 &  -1 & -1 & -1 & 1 & 1 & 1 & 1
& 9/4\\
\hline
$\chi^2 \rho c_{ix}$ & 0 &  1 &  -1 &  0&  0 &  0 &  0 & 
         0 &  0 &  0 &  0 &  -1 &  -1 & 1 & 1 & 0 & 0 & 0 & 0
& 9/2\\
\hline
$\chi^2 \rho c_{iy}$ & 0 &  0 &  0 & -1&   1 &  0 &  0 & 
         0 &  0 &  0 &  0 &   0 &  0 & 0 & 0 & 1 &  1 & -1 & -1
& 9/2\\
\hline
$\chi^2 \rho c_{iz}$ & 0 &  0 &  0 &  0&  0 &  0 &  0 & 
         0 &  0 &  0 &  0 &  -1 &  1 & -1 & 1 & 1 & -1 & 1 & -1
& 9/2\\
\hline
$\chi^3$ & 1 &  -2 &  -2 &  -2&  -2 &  -2 &  -2 & 
         1 &  1 &  1 &  1 &  1 &  1 & 1 & 1 & 1 & 1 & 1 & 1
& 1/2\\
\hline\hline
$w_i$ & 12 & 2 & 2 & 2 & 2 & 2 & 2 & 
1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1
& \\
\hline\hline
\end{tabular}
\caption{Table showing the details of the basis used for the D3Q19 model in
three dimensions. The nineteen modes $M^a$ include ten hydrodynamic modes,
three scalr kinetic modes $\chi^1$, $\chi^2$, and $\chi^3$; there are also
two vector kinetic modes $\chi^1 \rho c_{i\alpha}$
and $\chi^2 \rho c_{i\alpha}$. The weights in the equilibrium distribution
function are $w_i$, and the normaliser for each mode is $N^a$. The
eigenvectors of the collision matrix are the columns of the transformation
matrix $m^a_i$.
\label{table-d3q19-spec}
}
\end{table}

@@

For the conserved modes $\xi_i = 0$. For all the non-conserved modes,
i.e., those with dissipation, the fluctuating part may be written
\begin{equation}
\xi_i (\mathbf{r}; t) = w_i m_i^a \hat{\xi}^a (\mathbf{r}; t) N^a
\end{equation}
where $\hat{\xi}^a$ is a noise termwhich has a variance determined
by the relaxation time for given mode
\begin{equation}
\left< \hat{\xi}^a \hat{\xi}^b \right> =
\frac{\tau_a + \tau_b + 1}{\tau_a \tau_b}
\left< \delta M^a \delta M^b \right>.
\label{eq_fvar}
\end{equation}

@@

For the stress, the random contribution to the distributions is
\begin{equation}
\xi_i = w_i \frac{Q_{i\alpha\beta} \hat{\sigma}_{\alpha\beta}}{4c_s^2}
\end{equation}
where $\hat{\sigma}_{\alpha\beta}$ is a symmtric matrix of random
variates drawn from a Gaussian distribution with variance given
by equation~\ref{eq_fvar}. In the case that the shear and bulk
viscosities are the same, i.e., there is a single relaxation
time, then the variances of the six independent components of
the matrix are given by
\begin{equation}
\left< \hat{\sigma}_{\alpha\beta} \hat{\sigma}_{\mu\nu} \right> =
\frac{2\tau + 1}{\tau^2}
(\delta_{\alpha\mu}\delta_{\beta\nu} + \delta_{\alpha\nu} \delta_{\beta\mu}).
\end{equation}

>>>>

We review here the lattice Boltzmann method applied to a simple
Newtonian fluid with particular emphasis on the relevant
implementation in \textit{Ludwig}.

@@

In lattice Boltzmann, the density and velocity of the continuum fluid
are complemented by the  distribution function
$f_i(\mathbf{r}; t)$ defined with reference to the
discrete velocity space $c_{i\alpha}$.
It is possible to relate the hydrodynamic quantities to the distribution
function via its moments, that is
\begin{equation}
\rho(\mathbf{r};t) = \sum_i f_i(\mathbf{r};t),  \quad
\rho u_\alpha(\mathbf{r};t) = \sum_i f_i(\mathbf{r};t) c_{i\alpha},  \quad
\Pi_{\alpha\beta}(\mathbf{r};t) =
\sum_i f_i(\mathbf{r};t) c_{i\alpha} c_{i\beta}.
\label{equation-lb-f-moments}
\end{equation}
Here, the index of the summation is over the number of discrete
velocities used as the basis, a number which will be denoted
$N_\mathrm{vel}$. For example, in three dimensions
$N_\mathrm{vel}$ is often 19 and the basis is referred to as D3Q19.

The number of moments, or modes, supported by a velocity
set is exactly $N_\mathrm{vel}$, and these can be written in general as
\begin{equation}
M^a(\mathbf{r};t) = \sum_i m_i^a f_i(\mathbf{r};t),
\end{equation}
where the $m_i$ are the eigenvectors of the collision matrix in the LBE.
For example, in the case of the density, all the 
$m_i^a = 1$ and the mode $M^a$ is the density $\rho = \sum_i f_i$. Note
that the number of modes supported by a given basis will generally exceed the
number of hydrodynamic modes; the excess modes have no direct physical
interpretation and are variously referred to as non-hydrodynamic, kinetic,
or ghost, modes.
The ghost modes take no part in bulk hydrodynamics, but may become important
in other contexts, such as thermal fluctuations and near boundaries.
The distribution function can be related to the modes
$M^a(\mathbf{r};t)$ via
\begin{equation}
f_i(\mathbf{r};t) = w_i \sum_a m_i^a N^a M^a(\mathbf{r};t).
\end{equation}
In this equation, $w_i$ are the standard LB weights appearing in the
equilibrium distribution function, while the $N^a$ are a per-mode
normalising factor uniquely determined by the orthogonality condition
\begin{equation}
N^a \sum_i w_i m_i^a m_i^b = \delta_{ab}.
\end{equation}
Writing the basis this way has the advantage that the equilibrium
distribution projects directly into the hydrodynamic modes only.
Putting it another way, we may write
\begin{equation}
f_i^\mathrm{eq} = w_i \big(\rho + \rho c_{i\alpha}u_\alpha / c_s^2
+ (c_{i\alpha} c_{i\beta} - c_s^2\delta_{\alpha\beta})
(\Pi_{\alpha\beta}^\mathrm{eq} - p\delta_{\alpha\beta})/2c_s^4 \big)
\end{equation}
where only (equilibrium) hydrodynamic quantities appear on the right hand side.

@@

\begin{table}[t]
\begin{center}
\begin{tabular}{|l|r|rrrrrrrrr|r|l|}
\hline\hline
$M^a$ & $p$ & \multicolumn{9}{c|}{$m_i^a$} & $N^a$  &\\
\hline
$\rho$ & - & 1 &  1 &  1 &  1 &  1 &  1 &  1 &   1 &  1 & 1 &$\mathbf{1}$ \\
\hline
$\rho c_{ix}$ & - & 0 &  1 &  1 & 1 & 0 &  0 & -1 &  1 & -1 & 3 & $c_{ix}$ \\
\hline
$\rho c_{iy}$ & - & 0 & 1 &  0 &  -1 &  1 &  -1 & 1 & 0 & -1 & 3  &$c_{iy}$ \\
\hline
$Q_{xx}$ & 1/3 & -1 &  2 &  2 & 2 & -1 & -1 & 2 & 2 & 2 & 9/2 
& $c_{ix} c_{ix} - c_s^2$ \\
\hline
$Q_{xy}$ & - & 0 &  1 & 0 & -1 & 0 & 0 & -1 & 0 & 1 & 9 & $c_{ix} c_{iy}$ \\
\hline
$Q_{yy}$ & 1/3 & -1 &  2 & -1 & 2 & 2 & 2 & 2 & -1 & 2 & 9/2
& $c_{iy} c_{iy} - c_s^2$ \\
\hline\hline
$\chi^1$ & - &  1 & 4 & -2 & 4 & -2 & -2 & 4 & -2 & 4 & 1/4 & $\chi^1$ \\
\hline
$J_{ix}$ & - & 0 &  4 & -2 & 4 & 0 & 0 & -4 & -2 & -4 & 3/8
& $\chi^1 \rho c_{ix}$\\
\hline
$J_{iy}$ & - & 0 & 4 & 0 & -4 & -2 & 2 & 4 & 0 & -4 & 3/8
& $\chi^1 \rho c_{iy}$\\
\hline\hline
$w_i$ & 1/36 & 16 & 1 & 4 & 1 & 4 & 4 & 1 & 4 & 1 & & $w_i$\\
\hline\hline
\end{tabular}
\end{center}
\caption{Table showing the details of the basis used for the D2Q9 model
in two dimensions. The nine modes $M^a$ include six hydrodynamic modes,
one scalar kinetic mode $\chi^1$, and one vector kinetic mode $J_{i\alpha}$.
The weights in the equilibrium distribution function are $w_i$ and the
normaliser for each mode is $N^a$. The eigenvectors of the collision
matrix are the columns of the transformation matrix $m^a_i$. The pre-factor
$p$ (where present) multiplies all the elements to the right in that row.}
\label{table-d2q9-spec}
\end{table}

@@

\caption{Table showing the details of the basis used for the D3Q15 model
in three dimensions. The fifteen modes $M^a$ include two scalar kinetic
modes $\chi^1$ and $\chi^3$, and one vector kinetic mode $J_{i\alpha}$.
The weights in the equilibrium distribution are $w_i$ and the normaliser
for each mode is $N^a$. The eigenvectors of the collision matrix are the
columns of the transformation matrix $m_i^a$. The pre-factor $p$ simply
multiplies all elements of $m_i^a$ in that row as a convenience.
\label{table-d3q15-spec}
}
\end{table}

@@

\begin{table}[t]
\centering
\tabcolsep=4pt
\begin{tabular}{|l||r|rrrrrr|rrrr|rrrr|rrrr|r||}
\hline\hline
$M^a$ & \multicolumn{19}{c||}{$m_i^a$} & $N^a$\\
\hline
$\rho $ & 1 &  1 &  1 &  1 &  1 &  1 &  1 & 
         1 &  1 &  1 &   1 &  1 &  1 &  1 & 1 & 1 & 1 & 1 & 1
& 1\\
\hline
$\rho c_{ix}$ & 0 &  1 &  -1 &  0 &  0 &  0 &  0 & 
         1 &  1 &  -1 &   -1 &  1 &  1 &  -1 & -1 & 0 & 0 & 0 & 0
& 3 \\
\hline
$\rho c_{iy}$ & 0 &  0 &  0 &  1 &  -1 &  0 &  0 & 
         1 &  -1 &  1 &   -1 &  0 &  0 &  0 & 0 & 1 & 1 & -1 & -1
& 3\\
\hline
$\rho c_{iz}$ & 0 &  0 &  0 &  0 &  0 &  1 &  -1 & 
         0 &  0 &  0 &   0 &  1 &  -1 &  1 & -1 & 1 & -1 & 1 & -1
& 3\\
\hline
$Q_{ixx}$ & -1 &  2 &  2 &  -1&  -1 &  -1 &  -1 & 
         2 &  2 &  2 &   2 &  2 &  2 &  2 & 2 & -1 & -1 & -1 & -1
& 9/2\\
\hline
$Q_{iyy}$ & -1 &  -1 &  -1 &  2&  2 &  -1 &  -1 & 
         2 &  2 &  2 &   2 &  -1 &  -1 &  -1 & -1 & 2 & 2 & 2 & 2
& 9/2\\
\hline
$Q_{izz}$ & -1 &  -1 &  -1 &  -1&  -1 &  2 &  2 & 
         -1 &  -1 &  -1 &   -1 &  2 &  2 & 2 & 2 & 2 & 2 & 2 & 2
& 9/2\\
\hline
$Q_{ixy}$ & 0 &  0 &  0 &  0&  0 &  0 &  0 & 
          1 &  -1 &  -1 &    1 &  0 &  0 & 0 & 0 & 0 & 0 & 0 & 0
& 9\\
\hline
$Q_{ixz}$ & 0 &  0 &  0 &  0&  0 &  0 &  0 & 
          0 &   0 &   0 &   0 &  1 & -1 & -1 & 1 & 0 & 0 & 0 & 0
& 9\\
\hline
$Q_{iyz}$ & 0 &  0 &  0 &  0&  0 &  0 &  0 & 
          0 &   0 &   0 &   0 &  0 & 0 & 0 & 0 & 1 & -1 & -1 & 1
& 9\\
\hline\hline
$\chi^1$ & 0 &  1 &  1 &  1 &  1 &  -2 &  -2 & 
         -2 &  -2 &  -2 &  -2 &  1 &  1 & 1 & 1 & 1 & 1 & 1 & 1
& 3/4\\
\hline
$\chi^1 \rho c_{ix}$ & 0 &  1 &  -1 &  0&  0 &  0 &  0 & 
         -2 &  -2 &  2 &  2 &  1 &  1 & -1 & -1 & 0 & 0 & 0 & 0
& 3/2\\
\hline
$\chi^1 \rho c_{iy}$ & 0 &  0 &  0 &  1&  -1 &  0 &  0 & 
         -2 &  2 &  -2 &  2 &  0 &  0 & 0 & 0 & 1 & 1 & -1 & -1
& 3/2\\
\hline
$\chi^1 \rho c_{iz}$ & 0 &  0 &  0 &  0&  0 &  -2 &  2 & 
         0 &  0 &  0 &  0 &  1 &  -1 & 1 & -1 & 1 & -1 & 1 & -1
& 3/2\\
\hline
$\chi^2$ & 0 &  1 &  1 &  -1&  -1 &  0 &  0 & 
         0 &  0 &  0 &  0 &  -1 &  -1 & -1 & -1 & 1 & 1 & 1 & 1
& 9/4\\
\hline
$\chi^2 \rho c_{ix}$ & 0 &  1 &  -1 &  0&  0 &  0 &  0 & 
         0 &  0 &  0 &  0 &  -1 &  -1 & 1 & 1 & 0 & 0 & 0 & 0
& 9/2\\
\hline
$\chi^2 \rho c_{iy}$ & 0 &  0 &  0 & -1&   1 &  0 &  0 & 
         0 &  0 &  0 &  0 &   0 &  0 & 0 & 0 & 1 &  1 & -1 & -1
& 9/2\\
\hline
$\chi^2 \rho c_{iz}$ & 0 &  0 &  0 &  0&  0 &  0 &  0 & 
         0 &  0 &  0 &  0 &  -1 &  1 & -1 & 1 & 1 & -1 & 1 & -1
& 9/2\\
\hline
$\chi^3$ & 1 &  -2 &  -2 &  -2&  -2 &  -2 &  -2 & 
         1 &  1 &  1 &  1 &  1 &  1 & 1 & 1 & 1 & 1 & 1 & 1
& 1/2\\
\hline\hline
$w_i$ & 12 & 2 & 2 & 2 & 2 & 2 & 2 & 
1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1
& \\
\hline\hline
\end{tabular}
\caption{Table showing the details of the basis used for the D3Q19 model in
three dimensions. The nineteen modes $M^a$ include ten hydrodynamic modes,
three scalar kinetic modes $\chi^1$, $\chi^2$, and $\chi^3$; there are also
two vector kinetic modes $\chi^1 \rho c_{i\alpha}$
and $\chi^2 \rho c_{i\alpha}$. The weights in the equilibrium distribution
function are $w_i$, and the normaliser for each mode is $N^a$. The
eigenvectors of the collision matrix are the columns of the transformation
matrix $m^a_i$.
\label{table-d3q19-spec}
}
\end{table}

@@

For the conserved modes $\xi_i = 0$. For all the non-conserved modes,
i.e., those with dissipation, the fluctuating part may be written
\begin{equation}
\xi_i (\mathbf{r}; t) = w_i m_i^a \hat{\xi}^a (\mathbf{r}; t) N^a
\end{equation}
where $\hat{\xi}^a$ is a noise term which has a variance determined
by the relaxation time for given mode
\begin{equation}
\left< \hat{\xi}^a \hat{\xi}^b \right> =
\frac{\tau_a + \tau_b + 1}{\tau_a \tau_b}
\left< \delta M^a \delta M^b \right>.
\label{eq_fvar}
\end{equation}

@@

For the stress, the random contribution to the distributions is
\begin{equation}
\xi_i = w_i \frac{Q_{i\alpha\beta} \hat{\sigma}_{\alpha\beta}}{4c_s^2}
\end{equation}
where $\hat{\sigma}_{\alpha\beta}$ is a symmetric matrix of random
variates drawn from a Gaussian distribution with variance given
by equation~\ref{eq_fvar}. In the case that the shear and bulk
viscosities are the same, i.e., there is a single relaxation
time, then the variances of the six independent components of
the matrix are given by
\begin{equation}
\left< \hat{\sigma}_{\alpha\beta} \hat{\sigma}_{\mu\nu} \right> =
\frac{2\tau + 1}{\tau^2}
(\delta_{\alpha\mu}\delta_{\beta\nu} + \delta_{\alpha\nu} \delta_{\beta\mu}).
\end{equation}
